### 2019-04-23 Sections/intro_to_QIT.tex
Markup ZH's edits

\subsection{Probability, information and classical correlation measures}

``Information theory begins with the observation that there is a fundamental link between probabilities and information." With any given event (we call this a random variable), associated with such an event is
a set of outcomes, each with a certain probability of occuring.

Suppose we have a random variable $X$, and a particular measurement outcome $x$ occurs with probability $p(x)$, the Shannon entropy of $X$ is given by,
\begin{align}\index{Shannon entropy}
H(X) = -\sum_x p_x\log_2(p_x).
\end{align}

@@

 For a joint distribution over $X$ and $Y$ this generalises to,
\begin{align}\index{Joint Shannon entropy}
H(X,Y) =  -\sum_{x,y} p_{x,y}\log_2(p_{x,y}).
\end{align}

\begin{figure}[hbt]
\includegraphics[clip=true, width=0.45\textwidth]{classical_mutual_info.pdf}
\captionspacefig \caption{\label{fig:mutual_info}}	
\end{figure}

@@

For two classical distributions, the classical mutual information is given by,
\begin{align}\index{Mutual information}
I(A;B) = H(A) + H(B) - H(A,B).
\end{align}

It is a measure of the correlation between the events A and B. If these correspond, respectively, to the selection and receipts of a signal then H(A : B) is the information transferred by the communication.

The relationship between the quantities mentioned above is graphically summarised in Figure~\ref{fig:mutual_info}.

The von Neuman entropy\index{von Neuman entropy} \cite{bib:bengtsson2017geometry} for quantum density operators, $S(\hat\rho)$, is defined analogously, replacing probabilities with density operator eigenvalues,
\begin{align}\index{von Neuman entropy}
S(\hat\rho) &= - \sum_x \lambda_x \log_2 (\lambda_x) \nonumber \\
&= -\mathrm{tr}(\hat\rho\,\log \,\hat\rho),
\end{align}
where $\{\lambda\}$ is the eigenvalue spectrum of $\hat\rho$. This modification is logically justified, as the eigenvalues can be interpreted directly as a purely classical probability distribution of orthogonal states when the density operator is transformed into a basis with no coherences between basis states (i.e a diagonal basis or spectral decomposition\index{Spectral decomposition}). In that case, the Shannon and von Neuman entropies essentially have identical physical interpretations.

@@

A quantum process cannot increase the mutual information between two parties. This yields the \textit{data processing inequality}\index{Data processing inequality} that, for a sequence of channels \mbox{$X\to Y\to Z$},
\begin{align}\index{Data processing inequality}\label{eq:data_proc_ineq}
I(X:Z)&\leq I(X:Y), \nonumber \\
I(X:Z)&\leq I(Y:Z),
\end{align}
with equality if and only if the channel not specified in the identity on the right hand side (\mbox{$Y\to Z$} or \mbox{$X\to Y$} respectively) is unitary, i.e one of the links in the chain perfectly preserves information content. The progression is shown in Fig.~\ref{fig:data_proc_ineq}.

@@

% \paragraph{Quantum correlation measures}

@@

Analogous to the mutual information for classical systems is the \textit{coherent information} for quantum systems \cite{bib:PhysRevA.54.2629}, defined as,
\begin{align}\index{Coherent information}
I(\hat\rho,\mathcal{E}) = S(\mathcal{E}(\hat\rho)) - S_e(\hat\rho,\mathcal{E}).
\label{eq:mutual_info_classical_single}
\end{align}
Here $S_e$ is the \textit{exchange entropy}, a measure of how much information is exchanged between the state $\hat\rho$ and the environment under the action of the process. Specifically, it is given by the entropy of the environment subsystem in Eq.~(\ref{eq:proc_environment}), after application of the channel. This yields the intuitive interpretation that the coherent information is the information contained in the evolved state, discounted by the amount lost to the environment. It can be alternatively written as,
\begin{align}
I(A\rangle B) = H(B)_{\hat\rho} - H(A,B)_{\hat\rho},
\end{align}
which we recognise as the negative of the \textit{conditional quantum entropy}\index{Conditional quantum entropy},
\begin{align}
H(A|B) = H(A,B)_{\hat\rho} - H(B).
\end{align}
The fact that the information quantity $H(A|B)$ can be negative is a signature of the fundamental difference between the laws of classical and quantum information.

>>>>

% Probability, information & classical correlation measures

\subsection{Probability, information \& classical correlation measures}

``Information theory begins with the observation that there is a fundamental link between probabilities and information.''\comment{Is this a quote? Who?} With any given event (we call this a random variable), associated with such an event is a set of outcomes, each with a certain probability of occurring.

Suppose we have a random variable $X$, and a particular measurement outcome $x$ occurs with probability $p(x)$, the Shannon entropy\index{Shannon entropy} of $X$ is given by,
\begin{align}
	H(X) = -\sum_x p_x\log_2(p_x).
\end{align}

@@

For a joint distribution over $X$ and $Y$ this generalises to become the joint Shannon entropy\index{Joint Shannon entropy},
\begin{align}
H(X,Y) = -\sum_{x,y} p_{x,y}\log_2(p_{x,y}).
\end{align}

\begin{figure}[!hbtp]
	%\includegraphics[clip=true, width=0.475\textwidth]{classical_mutual_info}
	\captionspacefig \caption{\comment{Missing PDF. To do}}\label{fig:mutual_info}	
\end{figure}

\comment{Should we include Mike and Ike styled Venn diagram for information relationships?}

@@

For two classical distributions, the classical mutual information\index{Mutual information} is given by,
\begin{align}
	I(A;B) = H(A) + H(B) - H(A,B).
\end{align}

It is a measure of the correlation between the events $A$ and $B$. If these correspond, respectively, to the selection and receipts of a signal then \mbox{$H(A:B)$} is the information transferred by the communication.

The relationship between the quantities mentioned above is graphically summarised in Fig.~\ref{fig:mutual_info}.

% Quantum information

\subsection{Quantum information}

The von Neuman entropy\index{von Neuman entropy} \cite{bib:bengtsson2017geometry} for quantum density operators, $S(\hat\rho)$, is defined analogously, replacing probabilities with density operator eigenvalues,
\begin{align}
S(\hat\rho) &= - \sum_x \lambda_x \log_2 (\lambda_x) \nonumber \\
&= -\mathrm{tr}(\hat\rho\,\log \,\hat\rho),
\end{align}
where $\{\lambda\}$ is the eigenvalue spectrum of $\hat\rho$. This modification is logically justified, as the eigenvalues can be interpreted directly as a purely classical probability distribution of orthogonal states when the density operator is transformed into a basis with no coherences between basis states (i.e a diagonal basis or spectral decomposition\index{Spectral decomposition}). In that case, the Shannon and von Neuman entropies essentially have identical physical interpretations.

@@

A quantum process (the most general form of quantum evolution, described in Sec.~\ref{sec:quantum_processes}) cannot increase the mutual information between two parties. This yields the \textit{data processing inequality}\index{Data processing inequality} that, for a sequence of channels \mbox{$X\to Y\to Z$},
\begin{align}\index{Data processing inequality}\label{eq:data_proc_ineq}
	I(X:Z)&\leq I(X:Y), \nonumber \\
	I(X:Z)&\leq I(Y:Z),
\end{align}
with equality if and only if the channel not specified in the identity on the right hand side (\mbox{$Y\to Z$} or \mbox{$X\to Y$} respectively) is unitary, i.e one of the links in the chain perfectly preserves information content. The progression is shown in Fig.~\ref{fig:data_proc_ineq}.

@@

% Quantum correlation measures

@@

Analogous to the mutual information for classical systems is the \textit{coherent information}\index{Coherent information} for quantum systems \cite{bib:PhysRevA.54.2629}, defined as,
\begin{align}\label{eq:mutual_info_classical_single}
I(\hat\rho,\mathcal{E}) = S(\mathcal{E}(\hat\rho)) - S_e(\hat\rho,\mathcal{E}).
\end{align}
Here $S_e$ is the \textit{exchange entropy}, a measure of how much information is exchanged between the state $\hat\rho$ and the environment under the action of the process. Specifically, it is given by the entropy of the environment subsystem in Eq.~(\ref{eq:proc_environment}), after application of the channel. This yields the intuitive interpretation that the coherent information is the information contained in the evolved state, discounted by the amount lost to the environment. It can be alternatively written as,
\begin{align}
I(A\rangle B) = H(B)_{\hat\rho} - H(A,B)_{\hat\rho},
\end{align}
which we recognise as the negative of the \textit{conditional quantum entropy}\index{Conditional quantum entropy},
\begin{align}
H(A|B) = H(A,B)_{\hat\rho} - H(B).
\end{align}
The fact that the information quantity $H(A|B)$ can be negative is a signature of the fundamental difference between the laws of classical and quantum information.
